<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Introduction to httr}
-->

```{r, echo = FALSE}
library(httr)
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

# httr quickstart guide

The goal of this document is to get you up and running with httr as quickly as possible. It assumes some familiarity with the http protocol.

This vignette (and parts of the httr API) derived from the excellent "[Requests quickstart guide](http://docs.python-requests.org/en/latest/user/quickstart/)" by Kenneth Reitz.  

## HTTP verbs

To make a request, first load httr, then call `GET()` with a url:

```{r}
library(httr)
r <- GET("http://httpbin.org/get")
```

This gives you a response object. You can pull out important parts of the response with various helper methods, or dig directly into the object:

```{r}
http_status(r)
str(content(r))
str(r)
```

I'll use `httpbin.org` throughout this introduction. It accepts many types of http request and returns nicely formatted json describing the data that it recieved. This site is extremely useful if you want to check that httr is doing what you think it's doing.

As well as `GET()`, you can also use the `HEAD()`, `POST()`, `PATCH()`, `PUT()` and `DELETE()` verbs. You're probably most familiar with `GET()` and `POST()`: `GET()` is used by your browser when requesting a page, and `POST()` is used when submitting a form to a server. `PUT()`, `PATCH()` and `DELETE()` are used most often by web APIs.

## The response 

The data sent back from the server consists of two parts: the headers and the body. First I'll discuss the body, because that's often the data that you're most interested in.

### The body

There are three ways to access the body of the request, all using `content()`:

*   `content(r, "text")` accesses the body as text:

    ```{r}
    r <- GET("http://httpbin.org/get")
    content(r, "text")
    ```

    httr will automatically decode content from the server. When you make a 
    request, httr uses the encoding supplied in the HTTP headers. You can 
    override the defaults by supplying the encoding argument to `content()`: 

    ```{r}
    content(r, "text", encoding = "ISO-8859-1")
    ```

    Unfortunately websites don't always accurately declare the encoding that 
    they use. If you're having problems figuring out what the correct encoding 
    is, try `stringi::stri_enc_detect()`.

*   For non-text requests, you can access the body of the request as a 
    raw vector:

    ```{r}
    content(r, "raw")
    ```
    
    This is the safest way of saving files to disk:
    
    ```{r, eval = FALSE}
    bin <- content(r, "raw")
    writeBin("myfile.txt", bin)
    ```

*   httr provides a number of default parsers for common file types:

    ```{r}
    # JSON automatically parsed into named list
    str(content(r, "parsed"))
    ```
    
    These are convenient for interactive usage, but if you're writing an API
    wrapper, it's best to parse the text or raw content yourself. See the 
    API wrappers vignette for more details.

## Response headers

Access headers in the response with `headers()`:

```{r}
headers(r)
```

This is basically a named list, but because http headers are case insensitive, indexing ignores case:

```{r}
headers(r)$date
headers(r)$DATE
```

The most important headers are parsed and made available in other forms:

*   The status code:

    ```{r}
    r <- GET("http://httpbin.org/get")
    headers(r)$status
    http_status(r)
    ```

    (The best guide to status codes are the http status 
    [https://www.flickr.com/photos/girliemac/sets/72157628409467125](cats)).

    You can automatically throw a warning or raise an error if the request fails:
    
    ```{r}
    warn_for_status(r)
    stop_for_status(r)
    ```

*   The cookies:

    ```{r}
    r <- GET("http://httpbin.org/cookies/set", query = list(a = 1))
    cookies(r)
    ```
    
    Cookies are automatically persisted between requests to the same 
    domain:
    
    ```{r}
    r <- GET("http://httpbin.org/cookies/set", query = list(b = 1))
    cookies(r)
    ```

## Sending more data in the request

When sending a request to a server, there are a number of ways that you can send it more data:

* the query string in the url
* custom headers
* cookies
* the body of `POST()` requests

These are explained below.

### The url query string

A common way of sending simple key-value pairs to the server is the query string: e.g. `http://httpbin.org/get?key=val`. httr allows you to provide these arguments as a named list supplied to the `query` argument. For example, if you wanted to pass `key1=value1` and `key2=value2` to `http://httpbin.org/get` you could do:

```{r}
r <- GET("http://httpbin.org/get", 
  query = list(key1 = "value1", key2 = "value2"))
content(r)$args
```

Any `NULL` elements will be automatically dropped from the list, and both keys and values are escaped automatically.

```{r}
r <- GET("http://httpbin.org/get", 
  query = list(key1 = "value 1", "key 2" = "value2"))
content(r)$args
```

### Custom headers

You can add custom headers to a request with `add_headers()`:

```{r}
r <- GET("http://httpbin.org/get", add_headers(Name = "Hadley"))
str(content(r)$headers)
```

(Note that `content(r)$header` retrieves the headers that httpbin received. `headers(r)` gives the headers that it sent back in its response.)

## Cookies

Cookies are similar to url query strings (i.e. they're simple key-value pairs), but they persist across multiple requests in a session (because they're sent back and forth every time). To send your own cookies to the server, use `set_cookies()`:

```{r}
r <- GET("http://httpbin.org/cookies", set_cookies("MeWant" = "cookies"))
content(r)$cookies
```

Note that this response includes the `a` and `b` cookies that we created earlier.

### Post requests

When POSTing to a url, you normally include data in the `body` of the request. httr allows you to supply this in a number of different ways. The most common way is a named list:

```{r}
r <- POST("http://httpbin.org/post", body = list(a = 1, b = 2, c = 3))
```

You can use the `encode` argument to determine how this data is sent to the server:

```{r}
url <- "http://httpbin.org/post"
body <- list(a = 1, b = 2, c = 3)

# Form encoded
r <- POST(url, body = body, encode = "form")
# Multipart encoded
r <- POST(url, body = body, encode = "multipart")
# JSON encoded
r <- POST(url, body = body, encode = "json")
```

To see exactly what's being sent to the server, use `verbose()`:

```{r}
r <- POST(url, body = body, encode = "multipart", verbose()) # the default
r <- POST(url, body = body, encode = "form", verbose())
r <- POST(url, body = body, encode = "json", verbose())
```

`PUT()` and `PATCH()` work in identical ways.

## Posting files

There are two ways that you can send files:

```{r, eval = FALSE}
POST(url, body = upload_file("mypath.txt"))
POST(url, body = list(x = upload_file("mypath.txt")))
```

(`upload_file()` will guess the mime-type from the extension - using the `type` argument to override/supply yourself.)

Both of these uploads are streaming: the data will be loaded in R in chunks then sent to the remote server. This means that you can upload files that are larger than memory.

See `POST()` for more details on the other types of thing that you can send: no body, empty body, and character and raw vectors.

##### Built with

```{r}
sessionInfo()
```
