---
title: "Managing secrets"
author: "Hadley Wickham"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This document gives you the basics running on securely managing secrets. Relatively little is directly related to httr, but whenever you're working with an API you likely to have some secrets to handle. 

What is a secret? Some secrets are short alphanumeric sequences:

* Passwords (e.g. the second argument to `authenticate()`) are 
  clearly secrets. Passwords are particularly important because people
  often use the same password in multiple places.

* Personal access tokens used by websites like GitHub should be kept secret:
  they're equivalent to a user name password combination, but are slightly
  safer because you can have multiple tokens for different purposes and 
  it's easy to invalidate one token without affecting the others.

* Suprisingly, the "client secret" in an `oauth_app()` is not a secret.
  It's not equivalent to a password, and if you're are writing an API wrapper 
  package, it should be included in the package. (If you don't believe me,
  here are [google's comments on the topic][google-secret].)

Other secrets are files:

* The JSON web token (jwt) used for server-to-server OAuth 
  (e.g. [by google](https://developers.google.com/identity/protocols/OAuth2ServiceAccount)) 
  is a secret because it's equivalent to a personal access token.

* The `.httr-oauth` file is a secret because it stores OAuth access keys.

This vignette discusses the challenges of managing these secrets in locally

This document assumes that the main threat model is accidentally sharing your secrets when you don't want to. Protecting against a committed attacker is much harder and requires more set up at the operating system level (e.g. using full disk encryption) and is outside the scope of this document. Similarly, it's exceedingly difficult to protect secrets if someone can run code on your computer, and this is also out of scope.

Tradeoffs of convenience vs. security.

## Locally

It's usually fine to store secret files in the project directory as long as you take two precautions:

* Make sure they are listed in `.gitignore` so they don't accidentally 
  get included in a public repositry.
  
* Make sure they are listed in `.Rbuildignore` so they don't accidentally
  get included in a public R package.
  
The main remaining risk is that you zip up the entire directory to share it with someone else. If you're worried about this scenario, store your secret files outside of the project directory.

Storing short secrets is harder because it's tempting to record them as a variable as in your R script. This is a bad idea, and instead you either store in an environment variable, or use the keyring package as described below. 

### Ask every time

For rarely used scripts, one simple solution is to simply ask for the password each time. If you use RStudio an easy and secure way to request a password is:

```{r, eval = FALSE}
password <- rstudioapi::askForPassword()
```

You should __never__ type your password into the R console: this will typically be stored in the `.Rhistory` file, and it's easy to accidentally share without realising it.

For long or complicated passwords, it's fine to request the password in a regular text box:

```{r, eval = FALSE}
password <- rstudioapi::showPrompt("Password", "Your API password")
```

The only thing that the stars protect you from is someone shoulder surfing. This is generally not worth worrying about. It might be a concern if your user is working in an cafe, but even then if someone really wants to get their password they can video record your fingers and carefully reconstruct.

If you don't use RStudio, you can use a more general solution like the [getPass](https://github.com/wrathematics/getPass) package.

### Environment variables

Environment variables, or __env vars__ for short, are a cross platform way of storing strings outside of the running R process. Env vars are name value pairs stored in a file called `.Renviron` in your home directory. The easiest way to edit it is to run:

```{r, eval = FALSE}
file.edit("~/.Renviron")
```

The file looks something like

```
VAR1 = value1
VAR2 = value2
```

```{r, include = FALSE}
Sys.setenv("VAR1" = "value1")
```

```{r}
Sys.getenv("VAR1")
```

Note that `.Renviron` is only processed on startup, so you'll need to restart R for any changes to be reflected in your current session.

Note that these environment variables will be easily accessible by any R process. It's also possible for any other program running on your computer to access that file directly. You can get a little more security by using the system keyring.

### Keyring

The [keyring](https://github.com/r-lib/keyring) package provides a cross-platform way to store data in your operation system's secure secure store. 

It has a simple API:

```{r, eval = FALSE}
keyring::key_set("MY_SECRET")
keyring::key_get("MY_SECRET")
```

By default, keyring will use the system keyring which is unlocked. This means that while the password is stored securely pretty much running any process can access it. 

If you really want to be even more secure, you can custom keyring and lock it. Note that accessing the key always unlocks the keyring, so if you're being really careful, make sure to lock it again afterwards.

```{r, eval = FALSE}
keyring::keyring_create("httr")
keyring::key_set("MY_SECRET", keyring = "httr")
keyring::keyring_lock("httr")
keyring::key_get("MY_SECRET", keyring = "httr")
keyring::keyring_lock("httr")
```

## Sharing with others

By and large, managing secrets on your own computer is straightforward. The challenge comes when you need to share it with selected others. All of the approaches use __public key cryptography__. This is a type of assymetric encryption where you use a public key to produce content that can only be decrypted if you have the private key.

### Reprexes

The most common place you might need to share a secret is to generate a reprex. First, do everything you can do eliminate the need to share a secret:

* If it is an http problem, make sure to run all request with `verbose()`.
* If you get an R error, make sure to include `traceback()`.

If you're lucky, that will be sufficient information to fix the problem.

Otherwise, you'll need to encrypt the secret so you can share it with me. The easiest way to do so is with the following snippet:

```{r}
library(openssl)
library(jsonlite)
library(curl)

encrypt <- function(secret, username) {
  url <- paste("https://api.github.com/users", username, "keys", sep = "/")

  resp <- curl::curl_fetch_memory(url)
  pubkey <- jsonlite::fromJSON(rawToChar(resp$content))$key[1]

  opubkey <- openssl::read_pubkey(pubkey)
  cipher <- openssl::rsa_encrypt(charToRaw(secret), opubkey)
  jsonlite::base64_enc(cipher)
}
  
cipher <- encrypt("username,password", "hadley")
cipher
```

Then I can run the following code on my computer to access it:

```{r, eval = FALSE}
decrypt <- function(cipher, key = openssl::my_key()) {
  cipherraw <- jsonlite::base64_dec(cipher)
  rawToChar(openssl::rsa_decrypt(cipherraw, key = key))
}

decrypt(cipher)
#> username, password
```

I'll use the credentials carefully and won't share with anyone else, but I still recommend changing your password temporarily.

### GitHub

If you want to share secrets with a group of other people on GitHub, try the [secret](https://github.com/gaborcsardi/secret) package.

### Travis

Encrypted environment variables

https://docs.travis-ci.com/user/encryption-keys/

To run secret files on travis, see <https://docs.travis-ci.com/user/encrypting-files/>. Basically you will encrypt the file locally and check in to git. Then you'll add a decryption step to your `.travis.yml` which makes it decrypts it for each run.

Be careful to not accidentally expose the secret on travis. An easy way to accidentally expose the secret is to print it out so that it's captured in the log. Don't do that!

(Note that encrypted data is not available in pull requests in forks. Typically you'll need to check PRs locally once you've confirmed that the code isn't actively malicious.)

## CRAN 

There is no way to securely share information with arbitrary R users. This means that when developing a package you need to make sure `R CMD check` still passes without any problems when authentication is not available. The challenge is making sure this works while ensuring your code is still tested locally where authentication is available.

### Documentation

Like any R package, an API client needs clear and complete documentation of all functions. Examples are particularly useful but may need to be wrapped in `\donttest{}` to avoid challenges of authentication, rate limiting, lack of network access, or occasional API server down time.

### Vignettes

Vignettes pose additional challenges when an API requires authentication, because you don't want to bundle your own credentials with the package! However, you can take advantage of the fact that the vignette is built locally, and only checked by CRAN.

In a setup chunk, do:

```{r}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(purl = NOT_CRAN)
```

And then use `eval = NOT_CRAN` in any chunk that requires access to a secret.

### Testing

Use `testthat::skip_on_cran()` (preferrably wrapped inside a function like `auth_available()`) to skip all tests that require authentication.

[google-secret]: https://developers.google.com/identity/protocols/OAuth2#installed
